On the presentation of orthogonal groups over polynomial rings

### 2019-05-14 dl-paper.tex
Proved the last restricted axio

 The next step of the proof is to decompose each factor of the form $x_{\epsilon}(Xf)$, $\epsilon \in \{-\alpha,-\beta\}$ appearing in this product as
 $x_{\epsilon}(X^2f') \cdot x_{\epsilon}(mX)$ and then move $x_{\epsilon}(mX)$ to the rightmost position
  (conjugating by $x_{\epsilon}(mX)$ all factors along the way).
 Notice that no terms of the form $x_{\epsilon}(mX)$, $\epsilon \in \{ -\alpha, -\beta \}$ may appear after the simplification of conjugates,
  so this process is guaranteed to terminate.
 To see this notice that the extra terms in the right-hand sides of \eqref{eq3-1} and~\eqref{eq3-2} 
 (as well as all terms in the right-hand sides of~\eqref{C1},\eqref{C2} of~\cref{Crels} if one substitutes corresponding values into them)
  have terms of degree at least $2$ in $X$.

@@

\begin{lemma} The image of $K(\alpha, \beta)$ under $S_{\alpha}(a, -)$ is contained in the subgroup of $\St(\Phi, A[X, X^{-1}])$ 
 generated by $K(\alpha, \beta)$ and the subgroup $U$, which, by definition, is generated by $z_\gamma(f, X\xi)$, $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$. We call these elements generators of type 6.
\end{lemma}
\begin{proof} By~\eqref{eq:Zconj1}, \eqref{eq:Zconj2} the conjugate to each generator $z_\gamma(Xf, \xi)$ of types 1 by $x_\alpha(aX^{-1})$ is a product of generators of type 6, 6, 1.
By~\eqref{eq:Zconj3} the conjugate to the generator $x_{-\alpha}(X^2f)$ is a product of generators of type 1, 6, 1, 2, 1, 1, 6, 6.
Clearly, $S(\alpha, -)$ transforms $x_{-\beta}(X^2f)$ into the products of generators of type 2, 1.
Clearly, $S(\alpha, -)$ fixes both generators of type 3. 
As for the remaining generators type 5, \cref{Crels} shows that the corresponding conjugate can be written as a product of generators of type 6, 6, 5. \end{proof}

\begin{corollary} The image of $K(\alpha, \beta)$ under $S(\alpha, -)$ is contained in $P_\beta(0)$. \end{corollary}
\begin{proof} Follows from the above lemma and the fact that the generators of $U$ are generators of the second type for $P_\beta(0)$ in the sense of~\cref{defP0}.
 Indeed, if $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$ is such that $\langle \beta, \gamma \rangle = -1$, then $ \langle \alpha - \beta, \gamma \rangle \geq 2$, hence
  $\gamma = \alpha - \beta$, a contradiction. \end{proof}

\begin{lemma} Suppose $g \in K(A[X], M[X])$ then \[S_\alpha(a, g)\cdot  h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot m'a) \in P\left(\beta, \frac{m'}{1 + am}\right).\] \end{lemma}
\begin{proof}
 By~\cref{K-a-b} $g = g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $g_0\in K(\alpha, \beta)$.
 \begin{multline} \nonumber S_\alpha(a, g) = S_\alpha(a, g_0) \cdot S(a, x_{-\alpha}(mX)) \cdot S_\alpha\left(\frac{a}{1+am}, x_{-\beta}(m'X)\right) = \\ = S_\alpha(a, g_0) \cdot \langle -a, m \rangle \cdot h_\alpha(1+am) \cdot x_{-\beta}(m'X) \cdot x_{\alpha-\beta}\left(\frac{N_{\alpha, -\beta} \cdot m'a}{1+am}\right) = \\ =
  S_\alpha(a, g_0) \cdot \langle -a, m \rangle \cdot  x_{-\beta}\left(\frac{m'X}{1+am}\right) \cdot x_{\alpha-\beta}\left(N_{\alpha, -\beta} \cdot m'a\right) \cdot h_\alpha(1+am). \end{multline}
The assertion now follows from the preceding corollary.
\end{proof}

\begin{lemma} For $a \in A$ one has $K(\alpha, \beta)^{x_{\beta - \alpha}(a)} \subseteq K(\alpha, \beta)$. \end{lemma}
\begin{proof}
 By~\eqref{Z2}--\eqref{Z4} of~\cref{Zrels} the conjugate to the generator $g=z_\gamma(Xf, \xi)$ of type 1 by $x_{\beta-\alpha}(a)$ either coincides with $g$ or is a product of elements of type 4, 1, 1. Now if $g = x_\epsilon(Xf)$, $\epsilon \in \{\pm \alpha, \pm \beta\}$ is a generator of type 2 or 3, the conjugation by $x_{\beta-\alpha}(a)$ either fixes $g$ or transforms it into a product of either two elements of type 2 or two elements of type 3. By~\eqref{C1}--\eqref{C3} of~\cref{Crels} the conjugate of a generator $g = c_\delta(f, X\xi)$ of type 5 either coincides with $g$ or is a product of generators of type 5, 4, 1. Finally, from~\cref{rem:palpha} we obtain that $z_{\alpha-\beta}(Xf, a) \in K(A[X], M[X]) \cap P_\alpha(0) \cap P_{\beta}(0) = K(\alpha, \beta)$. \end{proof}

@@

Full axioms
\begin{enumerate}
 \item (+) $P(\alpha, m)$ contains $\mathrm{K}_2(\Phi, R[t, t^{-1}])$ (or at least the subgroup generated by symbols $\{ X, - \}$ and Steinberg/Dennis--Stein symbols);
 \item (+) if $g \in P(\alpha, m)$ then $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$;
 %\item if $\alpha \perp \beta$ then $g \in P(\alpha, m) \cap P(\beta, m')$ implies $S_\alpha(a, g) \in P(\beta, m')$;
 \item (+) if $\alpha + \beta \not \in \Phi$ but $\alpha - \beta \in \Phi$, $g \in P(\alpha, m) \cap P(\beta, m')$ then $S_\alpha(a, g)\cdot  h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-N_{\alpha-\beta, \beta}m'a) \in P(\beta, \frac{m'}{1 + am});$
 \item If $\alpha + \beta \in \Phi$ and $g \in P(\alpha, m) \cap P(\alpha + \beta, m')$ then
 $x_\beta(b) \cdot S_\alpha(a, g) \cdot h_{\beta}^{-1}(1+am)\cdot h_{\alpha}^{-1}(1 + N_{\alpha, \beta}abm')\cdot x_{\alpha}(-b(1 + N_{\alpha, \beta}abm')) \in P(\alpha, \frac{m - bm'}{1+am}).$
 \item if $\alpha + \beta \not \in \Phi$ then ???;
\end{enumerate}

Restricted axioms:
\begin{enumerate}
 \item (+) $P(\alpha, 0)$ contains $\mathrm{K}_2(\Phi, R[t, t^{-1}])$ (or at least the subgroup generated by symbols $\{ X, - \}$ and Steinberg/Dennis--Stein symbols);
 \item (+) if $g \in P(\alpha, 0)$ then $S_\alpha(a, g) \in P_\alpha(\alpha, 0)$;
% \item if $\alpha \perp \beta$ then $g \in P(\alpha, 0) \cap P(\beta, 0)$ implies $S_\alpha(a, g) \in P(\beta, 0)$;
 \item (+) if $\alpha + \beta \not \in \Phi$ but $\alpha - \beta \in \Phi$, $g \in P(\alpha, 0) \cap P(\beta, 0)$ then
$S_\alpha(a, g) \in P(\beta, 0);$
 \item If $\alpha + \beta \in \Phi$ and $g \in P(\alpha, 0) \cap P(\alpha + \beta, 0)$ then
 $x_\beta(b) \cdot S_\alpha(a, g) x_{\beta}(-b)\in P(\alpha, 0).$
 \item if $\alpha + \beta \not \in \Phi$ then ???;
\end{enumerate}

>>>>

 The next step of the proof is to decompose each factor of the form $x_{\epsilon}(Xf)$, $\epsilon \in \{-\alpha,-\beta\}$ appearing in this product as
 $x_{\epsilon}(X^2f') \cdot x_{\epsilon}(mX)$ and then move $x_{\epsilon}(mX)$ to the rightmost position
  conjugating by $x_{\epsilon}(mX)$ all factors along the way.
 Notice that no terms of the form $x_{\epsilon}(mX)$, $\epsilon \in \{ -\alpha, -\beta \}$ may appear after simplification of these conjugates, so this process is guaranteed to terminate.
 To see this notice that the extra terms in the right-hand sides of \eqref{eq3-1} and~\eqref{eq3-2} 
 (as well as all terms in the right-hand sides of~\eqref{C1},\eqref{C2} of~\cref{Crels} if one substitutes corresponding values into them)
  have terms of degree at least $2$ in $X$.

@@

 We need one more technical definition.
 \begin{df} Denote by $Z$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by elements $z_\gamma(f, X\xi)$, where $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$.
 \end{df}

\begin{lemma} \label{image-K-a-b} The image of $K(\alpha, \beta)$ under $S_{\alpha}(a, -)$ is contained in the subgroup of $\St(\Phi, A[X, X^{-1}])$  generated by $K(\alpha, \beta)$ and $Z$.
\end{lemma}
\begin{proof} 
To simplify notation we call the generators of the subgroup $Z$ ``generators of type 6``, despite the fact that these elements are not always contained in $K(\alpha, \beta)$ (or even $K(A[X], M[X])$).

By~\eqref{eq:Zconj1}, \eqref{eq:Zconj2} the conjugate to each generator $z_\gamma(Xf, \xi)$ of type 1 by $x_\alpha(aX^{-1})$ is a product of generators of type 6, 6, 1.
By~\eqref{eq:Zconj3} the conjugate to the generator $x_{-\alpha}(X^2f)$ is a product of generators of type 1, 6, 1, 2, 1, 1, 6, 6.
Clearly, $S(\alpha, -)$ transforms $x_{-\beta}(X^2f)$ into the products of generators of type 2, 1.
Clearly, $S(\alpha, -)$ fixes both generators of type 3. 
As for the remaining generators type 5, \cref{Crels} shows that the corresponding conjugate can be written as a product of generators of type 6, 6, 5. \end{proof}

\begin{corollary} The image of $K(\alpha, \beta)$ under $S(\alpha, -)$ is contained in $P_\beta(0)$. \end{corollary}
\begin{proof} Follows from the above lemma and the fact that the generators of $Z$ are generators of type 2 for $P_\beta(0)$ in the sense of~\cref{defP0}.
 Indeed, if $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$ is such that $\langle \beta, \gamma \rangle = -1$, then $ \langle \alpha - \beta, \gamma \rangle \geq 2$, hence
  $\gamma = \alpha - \beta$, a contradiction. \end{proof}

\begin{lemma} For $g \in K(A[X], M[X])$ there exist $m, m' \in M$ such that \[S_\alpha(a, g)\cdot  h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot m'a) \in P\left(\beta, \frac{m'}{1 + am}\right). \] \end{lemma}
\begin{proof}
 By~\cref{K-a-b} $g$ can be presented as $ g_0 \cdot x_{-\alpha}(mX) \cdot x_{-\beta}(m'X)$ for some $g_0\in K(\alpha, \beta)$, therefore from~\eqref{eq:Smult} and~\cref{lem:Tulenbayev-formula} we obtain that
 \begin{multline} \nonumber S_\alpha(a, g) = S_\alpha(a, g_0) \cdot S(a, x_{-\alpha}(mX)) \cdot S_\alpha\left(\frac{a}{1+am}, x_{-\beta}(m'X)\right) = \\ = S_\alpha(a, g_0) \cdot \langle -a, m \rangle \cdot h_\alpha(1+am) \cdot x_{-\beta}(m'X) \cdot x_{\alpha-\beta}\left(\frac{N_{\alpha, -\beta} \cdot m'a}{1+am}\right) = \\ =
  S_\alpha(a, g_0) \cdot \langle -a, m \rangle \cdot  x_{-\beta}\left(\frac{m'X}{1+am}\right) \cdot x_{\alpha-\beta}\left(N_{\alpha, -\beta} \cdot m'a\right) \cdot h_\alpha(1+am). \end{multline}
The assertion now follows from the preceding corollary.
\end{proof}

\begin{lemma} \label{conj-K-a-b} For $b \in A$ one has $K(\alpha, \beta)^{x_{\beta - \alpha}(b)} \subseteq K(\alpha, \beta)$. \end{lemma}
\begin{proof}
 By~\eqref{Z2}--\eqref{Z4} of~\cref{Zrels} the conjugate to the generator $g=z_\gamma(Xf, \xi)$ of type 1 by $x_{\beta-\alpha}(b)$ either coincides with $g$ or is a product of elements of type 4, 1, 1. Now if $g = x_\epsilon(Xf)$, $\epsilon \in \{\pm \alpha, \pm \beta\}$ is a generator of type 2 or 3, the conjugation by $x_{\beta-\alpha}(a)$ either fixes $g$ or transforms it into a product of either two elements of type 2 or two elements of type 3. By~\eqref{C1}--\eqref{C3} of~\cref{Crels} the conjugate to a generator $g = c_\delta(f, X\xi)$ of type 5 either coincides with $g$ or is a product of generators of type 5, 4, 1. Finally, from~\cref{rem:palpha} we obtain that $z_{\alpha-\beta}(Xf, b) \in K(A[X], M[X]) \cap P_\alpha(0) \cap P_{\beta}(0) = K(\alpha, \beta)$. \end{proof}

\begin{lemma} 
 For $g \in K(\alpha, \beta)$ and $a, b\in A$ the conjugate to
 $S_\alpha(a, g)$ by $x_{\beta-\alpha}(b)$ belongs to $P_\alpha(0)$.
\end{lemma}
\begin{proof}
 By~\cref{image-K-a-b} and~\cref{conj-K-a-b} it remains to show that the conjugate to each generator $g = z_\gamma(f, X\xi)$, $\gamma \in Z_+(\alpha) \setminus \{ \alpha - \beta \}$ of $Z$ by $x_{\beta-\alpha}(b)$ belongs to $P_\alpha(0)$.

 Throughout this proof we understand generator types according to~\cref{defP0}. It is clear that $g$ itself is a generator of type 2, therefore it remains to consider the case $\gamma\not \perp \beta - \alpha$.

 In the case $\alpha - \beta - \gamma \in \Phi$ we obtain from \eqref{Z2} of~\cref{Zrels} that
\[ z_{\gamma}(f, X\xi) ^ {x_{\beta-\alpha}(b)} = x_{\beta-\alpha} (- bXf\xi) \cdot x_{\gamma + \beta -\alpha} (N_{\gamma, \beta -\alpha}\cdot bf)     \cdot z_{\gamma}(f, X\xi)\ \text{if}\ \alpha - \beta - \gamma \in \Phi. \]
Notice that $\langle \gamma + \beta - \alpha, \alpha \rangle = 
  \langle \gamma , \alpha \rangle + \langle \beta , \alpha \rangle - \langle \alpha, \alpha \rangle \geq 1 + 1 - 2 = 0$, therefore the expression in the right-hand side is a product of generators of type 1, 2, 2 or 1, 3, 2. {\bf TODO: } Double check that both possibilities actually occur.

Now suppose that the other alternative holds, namely that $\alpha - \beta + \gamma \in \Phi$.  Since $\langle \beta - \gamma - \alpha , \alpha \rangle = \langle \beta , \alpha \rangle - \langle \gamma , \alpha \rangle - \langle \alpha, \alpha \rangle \leq 1 - 1 - 2 = -2$ we obtain that $\beta - \gamma - \alpha = -\alpha$. Thus, by \eqref{Z3} of~\cref{Zrels}
\[z_{\gamma}(f, X\xi) ^ {x_{\beta-\alpha}(b)} = x_{\beta-\alpha} (bXf\xi) \cdot x_{-\alpha} (N_{\gamma,\alpha - \beta}\cdot bX^2 f\xi^2) \cdot z_{\gamma}(f, X\xi). \]
The expression in the right-hand side is a product of generators of type 1, 4, 2. \end{proof}

@@

Full axioms
\begin{enumerate}
 \item (+) $P(\alpha, m)$ contains $\mathrm{K}_2(\Phi, R[t, t^{-1}])$ (or at least the subgroup generated by symbols $\{ X, - \}$ and Steinberg/Dennis--Stein symbols);
 \item (+) if $g \in P(\alpha, m)$ then $S_\alpha(a, g) \cdot h_\alpha^{-1}(1 + am) \in P(\alpha, \frac{m}{1 + am})$;
 %\item if $\alpha \perp \beta$ then $g \in P(\alpha, m) \cap P(\beta, m')$ implies $S_\alpha(a, g) \in P(\beta, m')$;
 \item (+) if $\alpha + \beta \not \in \Phi$ but $\alpha - \beta \in \Phi$, $g \in P(\alpha, m) \cap P(\beta, m')$ then $S_\alpha(a, g)\cdot  h^{-1}_\alpha(1 + am) \cdot x_{\alpha-\beta}(-N_{\alpha-\beta, \beta}m'a) \in P(\beta, \frac{m'}{1 + am});$
 \item If $\alpha + \beta \in \Phi$ and $g \in P(\alpha, m) \cap P(\alpha + \beta, m')$ then
 $x_\beta(b) \cdot S_\alpha(a, g) \cdot h_{\beta}^{-1}(1+am)\cdot h_{\alpha}^{-1}(1 + N_{\alpha, \beta}abm')\cdot x_{\alpha}(-b(1 + N_{\alpha, \beta}abm')) \in P(\alpha, \frac{m - bm'}{1+am}).$
\end{enumerate}

Restricted axioms:
\begin{enumerate}
 \item (+) $P(\alpha, 0)$ contains $\mathrm{K}_2(\Phi, R[t, t^{-1}])$ (or at least the subgroup generated by symbols $\{ X, - \}$ and Steinberg/Dennis--Stein symbols);
 \item (+) if $g \in P(\alpha, 0)$ then $S_\alpha(a, g) \in P_\alpha(\alpha, 0)$;
% \item if $\alpha \perp \beta$ then $g \in P(\alpha, 0) \cap P(\beta, 0)$ implies $S_\alpha(a, g) \in P(\beta, 0)$;
 \item (+) if $\alpha + \beta \not \in \Phi$ but $\alpha - \beta \in \Phi$, $g \in P(\alpha, 0) \cap P(\beta, 0)$ then
$S_\alpha(a, g) \in P(\beta, 0);$
 \item (+) If $\alpha + \beta \in \Phi$ and $g \in P(\alpha, 0) \cap P(\alpha + \beta, 0)$ then
 $x_\beta(b) \cdot S_\alpha(a, g) x_{\beta}(-b)\in P(\alpha, 0).$
\end{enumerate}
